Spatiotemporal Multi-Resolution Approximation of the Amari Type Neural Field Model

### 2012-08-21 Multiresolution_framework/ltx/response_to_reviews.tex

\section{Handling editor}
\begin{enumerate}
\item Whilst the paper does fit the scope of NI, I agree with R1 and R3, that explicit advantages of going to a multiresolution frame - in addition to the conceptual ones - compared to the prior work based on Gaussian basis functions needs clarification and elucidation.

\emph{\parham{The first paragraph of the newly added ``State-space representation'' subsection in the Discussion  highlights the benefits of the MRAIDE compared to the prior work. We also added a paragraph to `` Estimation algorithm'' in the Discussion to clarify the benefits of the novel M-step presented in this work compared to the implementation in Dewar et al.. Please refer to Responses 1 and 5 to R1 for more details.}}

\item  In addition, at least a ``proof of principle" inversion from empirical data would enhance the potential impact of the paper.

\emph{The authors agree that the use of empirical data will strengthen the paper. Unfortunately, the authors do not have access to the appropriate data at present. However, we understand that several other groups are now collecting high spatiotemporal resolution data that is appropriate for the methods developed in this paper. We believe that the methods developed in this paper are equally important as the development of high spatiotemporal resolution recording techniques and together, these approaches have enormous potential to answer some of the biggest questions in neuroscience. The ability to use mathematical models as filters, or lenses into the brain if you like, is still a relatively new concept to many members in our community. This new lens allows us to image aspects of neurophysiology of individual subjects that are normally hidden in electrical recordings. Therefore, we believe it is important to continue improve our methods that will lead to the next era in systems neuroscience, where mathematical modeling, experimental measures and model-data fusion will become standard methods. Furthermore, we believe that (good) experimental design is based around the availability of analysis tools. Therefore new imaging tools, like the framework presented in this paper, will lead to new data being collected and potential advances in our collective knowledge of brain function.}

@@

        \item Section 5 is a review of known results and standard background material from wavelet theory and includes a simple calculation of the Fourier transform of the particular basis used here, which would be more appropriate for an appendix (cf appendix H of the Freestone et al. paper where the Fourier transform of the Gaussian basis functions is calculated).

		% \emph{\parham{Section 5 summarizes important characteristics of B-Spline scaling and wavelet functions which are collected from different references, with an attempt to clarify the inner product and convolution of theses functions (please see R2 comment on ``B-spline calculus''). We think that this brief review can familiarize interested readers  with  B-spline scaling and wavelet functions and direct them to the important literature on this matter. As the reviewer suggested the calculation of the Fourier transform is now moved to the appendix.}}

\emph{As the reviewer suggested, moving the calculation of the Fourier transform to the appendix will provide a better flow to the paper. This has been done.}

\emph{We understand that this section will not be the central focus of the paper for some members of our research community, however we believe that it is important that this information remains in the paper to allow for implementation without the requirement of digesting information from several texts.}

\emph{We have had similar feedback to your suggestion from our colleagues prior to submission of the paper. Interestingly, the feedback from the people with expertise in wavelet theory thought that the overview of the estimation methods was essential, but the section on wavelets should be reduced. Conversely, people with expertise in the estimation methods thought the overview on wavelets was critical and the section on EM-algorithm should be reduced. We completely appreciate R1 comments and suggestions, but since R2 commented on the usefulness of this section we would very much like to leave it in. We believe that different sections will be valuable for people with different specializations, the synthesis of this information is a valuable contribution for implementation for a wide audience.}

@@

        \item Section 6 also borrows heavily from Dewar et al, e.g. cf equation (58) of this paper with equation (19) of Dewar et al: the state transition matrix is a perturbation of the corresponding matrix used in that earlier paper. The approach used in both papers seems to be the same with an RTS smoother used in the E-step (this was also used and explained in Freestone et al.) and an M-step optimizing a cost function, quadratic in the parameters, cf. equation (62) with equation (29) of Theorem 2 of Dewar et al. In general, the derivations of section 6 and appendix B follow closely or are very similar to those of Dewar et al.

		\emph{To clarify the advantage of the M-step over the implementation in Dewar et al. we have added a paragraph to Section 8.1 (now 8.2). Further comments for the reviewer are added below.}

\emph{The inverse framework in Dewar et al. is based on a general integro-differential equation model, which is not suitable when neural field models are considered. As the result the necessary additions have been made.}

\emph{This paper has a novel implementation of the M-step, which significantly simplifies the computational complexity of the EM-based algorithm proposed by Dewar et al.. These modifications are significant and crucial for high-dimensional systems. More specifically, implementing the multi-resolution framework introduced in the paper would be impossible without this new contribution. The differences are highlighted in Table.~4 of the manuscript, namely from $O(n_x^4n_{\theta}^2)$ to $O(n_x^2n_{\theta}^2)$, see equation (64) of this paper and equation (31) of Dewar et al, where the static and dynamic components are separated by defining $\boldsymbol \Gamma_1$ and $\mathbf U$ (see equations (65) and (67)). A simplification is also performed to calculate $\boldsymbol\upsilon_0, \boldsymbol\upsilon_1$ (see equations (62) and (63) of this paper and equation (30) of Dewar et al.). }

\emph{The use of the RTS smoother in the E-step of the EM algorithm for linear state-space models is an effective, standard and widely used approach. In fact, the aim of the IDE representation in a state-space form is to adopt the EM and consequently the RTS smoother in the E-step. The RTS algorithm is presented for completeness to interested readers.}

### 2012-08-21 Multiresolution_framework/ltx/response_to_reviews.tex

\emph{The use of the RTS smoother in the E-step of the EM algorithm for linear state-space models is an effective, standard and widely used approach. In fact, the aim of the IDE representation in a state-space form is to adopt the EM and consequently the RTS smoother in the E-step. The RTS algorithm is presented for completeness to interested readers.}

		\emph{\parham{For linear state-space models, under Gaussian conditions, the lower bound on the log-likelihood function is always quadratic in parameters, and can be solved analytically when the parameters are separable giving results equivalent to the least square estimates. In fact this is one of the advantages of this representation, such a closed form solution for the system's parameters results in significant computational savings.}}

        \item Also, Appendix A is the analogue of Appendix D of Freestone et al. and Appendix C includes results known from the literature (as the authors point out).

\emph{In order to formulate the state-space representation of the multi-resolution neural field model the inner product of the B-spline functions are required. We believe that it is necessary to provide the interested reader with the required tools to implement the algorithm. Appendix D of Freestone et al. and Appendix A herein are only similar in the sense that the inner product of two basis functions are calculated. In Appendix D of Freestone et al. two  n-dimensional Gaussians are considered, while here the inner product of two B-spline functions is calculated. The authors have no objections to remove Appendix C. However as stated above, we think that it would be informative for the broad audience of NeuroImage to include the model specific derivations and some of the well-known formulations of the EM algorithm, since it draws on a wide range of literature.}

@@

\emph{\parham{To respond to this important comment we have added ``State-space representation'' subsection into the discussion.}}

				\item{The framework of Freestone et al., seemed not to assume a linear relation between mean membrane potential and firing rate and in that sense the present approach appears more restrictive. }

\emph{\dean{The linear activation function in the estimation algorithm was made out of necessity. This was due to the high dimensionality and associated computational demands of the estimation problem.}}

\emph{\dean{The authors would like to point out that the derivation of the state-space multi-resolution model was made general for all possible activation functions. This derivation is perhaps the most important contribution of the paper \parham{not sure about this because we have already done the derivation in Freestone 2011}. We understand that all physical systems are nonlinear. However, we also understand that linear approximations are often required in engineering for useful models. In the introduction to the estimation section, we have attempted to discuss this issue with regard to forward vs. inverse modelling. \parham{Dean: I think something is missing here: We believe that the linear model is more acceptable for inverse modelling, but understand it is required for forward modelling.} For example, in a forward model the nonlinearity is required to prevent negative firing rates. However in the inverse case, the physiology naturally restricts firing rates to be positive. Furthermore, in the mean field case, we do not expect firing rates of all neurons to be zero or all neurons to be simultaneously firing at their maximal rate. Therefore, a linear model may suffice to image a multi-resolution neural field and physiological parameters using electrophysiological data .}}

### 2012-08-21 Multiresolution_framework/ltx/response_to_reviews.tex

				\item Also, the comments on large scale networks and delays made on p.20 look irrelevant to the present framework.

				\emph{\parham{We have omitted the last paragraph of the discussion on delays in large scale networks.}}

				\item Much of the discussion of the last section focuses on technical issues, while the discussion of the observation kernel chosen on p.17 is limited.

				\emph{\parham{We have added a paragraph in the new subsection ``State-space representation'' in the Discussion, regarding the choice of the observation kernel.}}

@@

			\emph{Response here.}

			\item Contrasting, for example, this multiresolution method with Freestone's 2011 results would be helpful for continuity.

			\emph{Response here.}

			% \emph{\parham{To respond to this comment we have added ``State-space representation'' subsection into the discussion.}}

### 2012-08-22 Multiresolution_framework/ltx/manuscript.tex
edits to letter and headings in paper

In summary, to solve the joint state and parameter estimation for linear state-space models one can adapt the EM algorithm, solving the smoothing problem for the E-step and constructing and maximizing the lower bound in the M-step. A schematic representation of the EM algorithm is represented in Fig.~\ref{fig:EMBlockDiagram}. Note the algorithm can be initialized by either a random parameter set or a sequence of the random bounded state vectors. Here the latter is chosen as shown in the block diagram ensuring the resultant initial kernel estimate leads to a stable system.  
\subsection{E-step}
As mentioned earlier, the E-step for state-space models is equivalent to the smoothing problem. We use the RTS smoother, given in Table~\ref{alg:MRA-RTS} for completeness, in order to infer the states of the system. The RTS smoother uses the Kalman filter \citep{Kalman1960} in forward pass (Steps~1 and 2 in Table~\ref{alg:MRA-RTS}), along with a backward iteration to calculate its outputs (Steps~3 and 4 in Table~\ref{alg:MRA-RTS}): smoothed states $\hat{\mathbf x}^b_t$ and covariances, $\mathbf P^b_t=\mathrm{cov}(\mathbf{x}_t)$. Following \cite{Gibsona2005}, an additional backwards recursion is also included (Step~5 in Table~\ref{alg:MRA-RTS})	to calculate cross-covariance matrix, $\mathbf M_t=\mathrm{cov}(\mathbf{x}_{t},\mathbf{x}_{t+1})$, which is also required in forming the lower bound on the likelihood function. Note that the following quantities from the forward iteration should be stored when the RTS smoother is implemented: predicted and corrected estimates for both state estimates ($\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^{f}$) and covariance matrices ($\mathbf P_t^{f-}$,$\mathbf P_t^f$). Also, the final value of the Kalman gain, $\mathcal K_T$, should be stored in order to initialize the backward cross-covariance computation. 
 \subsection{M-step}
The lower bound on the log-likelihood function is given by the expectation of the complete (hidden and observable) data  log-likelihood \citep{Bishop2006} 
\begin{equation}\label{eq:Bishopbound}
	\mathcal Q(\boldsymbol \theta,\boldsymbol\theta')= \mathbf E_{\boldsymbol \theta'}\left[2\ln p(\mathbf X,\mathbf Y;\boldsymbol \theta)\right], 
\end{equation}
where $p(\mathbf X,\mathbf Y;\boldsymbol \theta)$ is the joint probability distribution of the states and the observations and $ \mathbf E_{\boldsymbol \theta'}\left[\cdot\right] $ denotes expectation taken with respect to the marginal distribution of the states conditioned on the observed field and the current estimate of the parameter set, $\boldsymbol\theta'$ i.e.  $p(\mathbf X\mid\mathbf Y;\boldsymbol \theta'$). Note that $p(\mathbf X\mid\mathbf Y;\boldsymbol \theta'$), is in fact the smoother output which is obtained at the E-step.

@@

\subsection{Stopping rule}

@@

\subsection{Computational complexity} 

@@

One snap shot of the field reconstruction, at different levels of approximation is given in Fig.~\ref{fig:FieldEstimates100}, showing small discrepancy between the actual and estimated field at $j=3$. From this figure, it can be observed that the overall trend of the underlying field can be also captured at the coarser levels.
\subsection{Experiment II}
To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The mean over time of the root mean square error (MRMSE) over space of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ (used in the first experiment) is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution finite element model representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. By decomposing both the spatial field and the connectivity kernel using a wavelet decomposition, it becomes possible to represent small scale details in the field at the same time as large scale details. This work is significant, as the ability to create patient-specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling, where the dynamics and connectivity structure differs at different spatial scales~\citep{Jirsa2002,Jirsa2004,Breakspear2005,Qubbaj2009,Schultze-Kraft2011}. 
\parham{\subsection{State-space representation}
Different basis functions can be used to reduce the neural field formulations for state and parameter estimation problems. An alternative is given by the Gaussian basis functions adopted by \cite{Freestone2011} to decompose the field and connectivity kernel. One clear advantage of B-spline scaling and wavelet functions is the MRA capability, allowing the simultaneous representation of the neural field at different spatial scales. B-spline scaling and wavelet functions are compactly supported, eliminating  numerical errors caused by the Gaussian basis function. The partition of unity of these basis functions enable the reconstruction of the neural field when it saturates, this may be beneficial to model neurological disorders such as epilepsy. Attaining fine scale approximation of the neural field using Gaussian basis functions results into an ill-posed problem as the number of states is increased significantly, however, this is not the case when using B-spline scaling and wavelet functions due to their semi-orthogonality property.\\ \cite{Freestone2011} showed experimentally that Gaussian functions are a suitable choice for modeling the observation kernel when intracranial measurements are considered. B-splines converge to Gaussian functions as the order of the B-spline, $m$, tends to infinity. The cubic B-spline is already a very close approximation to the Gaussian function, therefore these should also be suitable for modeling the observation kernel while having the advantage of being compactly supported.}

\subsection{Estimation algorithm}
\parham{The implementation of the M-step in this work significantly simplifies the computational complexity of the EM based algorithm proposed by \cite{Dewar2009} as it is shown in Table~\ref{table:MRA-ComputationalComplexity}, namely from $O(n_x^4n_{\theta}^2)$ to $O(n_x^2n_{\theta}^2)$. This is crucial for high dimensional systems and more specifically for implementing the multi-resolution framework introduced herein which otherwise would be possible.}

@@

\parham{The basis structure (i.e., non-redundant, versus a (tight) frame representation) is a requirement for this framework.                                              }Choosing isotropic scaling and wavelet functions allows switching the order of operations with the inner product and convolution in Eq.~\eqref{eq:ApproxDiscreteTimeModel4}. This simplifies the computational complexity of the estimation algorithm significantly. Although this is correct for one dimensional B-spline scaling and wavelet functions, it does not hold at higher dimensions as the resulting B-spline wavelets are not isotropic. One alternative could be  
isotropic polyharmonic B-splines scaling functions and wavelets \citep{VanDeVille2005}.
 \subsection{Modification to higher dimensions}
 The reconstruction of the neural field in higher dimensions requires $2^n-1$ wavelets, where $n$ is the dimension of the space. For example a two-dimensional MRA can be implemented using 2-D scaling and wavelet functions built up using the tensor-product approach \citep{Meyer1992}. In this case three wavelet functions are required to  extract fine features of the field at vertical, horizontal and diagonal orientations. The formulations provided in this paper can be easily extended to two dimensions. However, the computational load of the estimation algorithm will be significant. It is important to investigate methods for choosing basis functions to balance the complexity with the reconstruction capability of the model. \parham{It is also possible to apply other than dyadic subsampling schemes in multiple dimensions, which leads to less wavelet subbands. For example, in 2D, the quincunx scheme is well-known to have better isotropic properties. Unfortunately, in 3D, it is not possible to have a subsampling operator (acting on a Cartesian sampling lattice) that is performing subsampling at equal rate in each dimension at each iteration \citep{VanDeVille2005a}.}
\subsection{Extensions to the framework} 
Clearly there is, and will always be, a discrepancy between the dynamics of the model and the experimental observation. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The MRAIDE approach is not limited to neural fields; the framework can be applied to modeling other multi-resolution spatiotemporal dynamical systems such as weather systems, ecological systems, and others~\citep{Wikle2002,Xu2005}. The key development is the multi-resolution decomposition forming the state-space model. While the decomposition holds for more sophisticated models, efficiently performing nonlinear smoothing in the large state-space remains a challenge.

>>>>

In summary, to solve the joint state and parameter estimation for linear state-space models one can adapt the EM algorithm, solving the smoothing problem for the E-step and constructing and maximizing the lower bound in the M-step. A schematic representation of the EM algorithm is represented in Fig.~\ref{fig:EMBlockDiagram}. Note the algorithm can be initialized by either a random parameter set or a sequence of the random bounded state vectors. Here the latter is chosen as shown in the block diagram ensuring the resultant initial kernel estimate leads to a stable system.  
\subsection{E-Step}
As mentioned earlier, the E-step for state-space models is equivalent to the smoothing problem. We use the RTS smoother, given in Table~\ref{alg:MRA-RTS} for completeness, in order to infer the states of the system. The RTS smoother uses the Kalman filter \citep{Kalman1960} in forward pass (Steps~1 and 2 in Table~\ref{alg:MRA-RTS}), along with a backward iteration to calculate its outputs (Steps~3 and 4 in Table~\ref{alg:MRA-RTS}): smoothed states $\hat{\mathbf x}^b_t$ and covariances, $\mathbf P^b_t=\mathrm{cov}(\mathbf{x}_t)$. Following \cite{Gibsona2005}, an additional backwards recursion is also included (Step~5 in Table~\ref{alg:MRA-RTS})	to calculate cross-covariance matrix, $\mathbf M_t=\mathrm{cov}(\mathbf{x}_{t},\mathbf{x}_{t+1})$, which is also required in forming the lower bound on the likelihood function. Note that the following quantities from the forward iteration should be stored when the RTS smoother is implemented: predicted and corrected estimates for both state estimates ($\hat{\mathbf{x}}_t^{f-}$, $\hat{\mathbf{x}}_t^{f}$) and covariance matrices ($\mathbf P_t^{f-}$,$\mathbf P_t^f$). Also, the final value of the Kalman gain, $\mathcal K_T$, should be stored in order to initialize the backward cross-covariance computation. 
 \subsection{M-Step}
The lower bound on the log-likelihood function is given by the expectation of the complete (hidden and observable) data  log-likelihood \citep{Bishop2006} 
\begin{equation}\label{eq:Bishopbound}
	\mathcal Q(\boldsymbol \theta,\boldsymbol\theta')= \mathbf E_{\boldsymbol \theta'}\left[2\ln p(\mathbf X,\mathbf Y;\boldsymbol \theta)\right], 
\end{equation}
where $p(\mathbf X,\mathbf Y;\boldsymbol \theta)$ is the joint probability distribution of the states and the observations and $ \mathbf E_{\boldsymbol \theta'}\left[\cdot\right] $ denotes expectation taken with respect to the marginal distribution of the states conditioned on the observed field and the current estimate of the parameter set, $\boldsymbol\theta'$ i.e.  $p(\mathbf X\mid\mathbf Y;\boldsymbol \theta'$). Note that $p(\mathbf X\mid\mathbf Y;\boldsymbol \theta'$), is in fact the smoother output which is obtained at the E-step.

@@

\subsection{Stopping Rule}

@@

\subsection{Computational Complexity} 

@@

One snap shot of the field reconstruction, at different levels of approximation is given in Fig.~\ref{fig:FieldEstimates100}, showing small discrepancy between the actual and estimated field at $j=3$. From this figure, it can be observed that the overall trend of the underlying field can be also captured at the coarser levels.
\subsection{Experiment II}
To study the performance of different models with various MRA capabilities, a single realization was obtained using Eqs.~\eqref{eq:DiscreteTimeModel} and \eqref{eq:ObservationEquation} under the same circumstances of the previous experiment. The generated data set was then employed to estimate the underlying field and the connectivity kernel parameters using different state-space models, $j=0$ up to $j=4$. The mean over time of the root mean square error (MRMSE) over space of the field estimation for different models is shown in Fig.~\ref{fig:RMSE}. By increasing the value of $j$, error in the estimation decreases and converges to a steady value of 0.71, confirming that the model with $j=3$ (used in the first experiment) is in fact adequate to capture the dynamics of the underlying field. The spatiotemporal characteristics of the reconstructed fields (A-E) along with the actual field (F)  are shown in Fig.~\ref{fig:FieldEstimation}. The experiment was performed several times and the results were consistent over each run.
\section{Discussion}
In this paper, we have presented a novel model-based framework for estimating cortical dynamics from electrophysiological measurements. The novel and key developments of the paper include the multi-resolution finite element model representation of the neural field, and the estimator for an intracortical connectivity kernel with an arbitrary shape. By decomposing both the spatial field and the connectivity kernel using a wavelet decomposition, it becomes possible to represent small scale details in the field at the same time as large scale details. This work is significant, as the ability to create patient-specific neural field models has the potential to contribute to our understanding and improve treatment of diseases resulting from abnormal neurodynamics. Other groups have previously highlighted the importance of a multi-resolution approach in neural field modeling, where the dynamics and connectivity structure differs at different spatial scales~\citep{Jirsa2002,Jirsa2004,Breakspear2005,Qubbaj2009,Schultze-Kraft2011}.

\parham{\subsection{State-Space Representation}
Different basis functions can be used to reduce the neural field formulations for state and parameter estimation problems. An alternative is given by the Gaussian basis functions adopted by \cite{Freestone2011} to decompose the field and connectivity kernel. One clear advantage of B-spline scaling and wavelet functions is the MRA capability, allowing the simultaneous representation of the neural field at different spatial scales. B-spline scaling and wavelet functions are compactly supported, eliminating  numerical errors caused by the Gaussian basis function. The partition of unity of these basis functions enable the reconstruction of the neural field when it saturates, this may be beneficial to model neurological disorders such as epilepsy. Attaining fine scale approximation of the neural field using Gaussian basis functions results into an ill-posed problem as the number of states is increased significantly, however, this is not the case when using B-spline scaling and wavelet functions due to their semi-orthogonality property.\\ \cite{Freestone2011} showed experimentally that Gaussian functions are a suitable choice for modeling the observation kernel when intracranial measurements are considered. B-splines converge to Gaussian functions as the order of the B-spline, $m$, tends to infinity. The cubic B-spline is already a very close approximation to the Gaussian function, therefore these should also be suitable for modeling the observation kernel while having the advantage of being compactly supported.}

\subsection{Estimation Algorithm}
\parham{The implementation of the M-step in this work significantly simplifies the computational complexity of the EM based algorithm proposed by \cite{Dewar2009} as it is shown in Table~\ref{table:MRA-ComputationalComplexity}, namely from $O(n_x^4n_{\theta}^2)$ to $O(n_x^2n_{\theta}^2)$. This is crucial for high dimensional systems and more specifically for implementing the multi-resolution framework introduced herein which otherwise would be possible.}

@@

\parham{The basis structure (i.e., non-redundant, versus a (tight) frame representation) is a requirement for this framework.                                              }Choosing isotropic scaling and wavelet functions allows switching the order of operations with the inner product and convolution in Eq.~\eqref{eq:ApproxDiscreteTimeModel4}. This simplifies the computational complexity of the estimation algorithm significantly. Although this is correct for one dimensional B-spline scaling and wavelet functions, it does not hold at higher dimensions as the resulting B-spline wavelets are not isotropic. One alternative could be  
isotropic polyharmonic B-splines scaling functions and wavelets \citep{VanDeVille2005}.

 \subsection{Modification to Higher Dimensions}
 The reconstruction of the neural field in higher dimensions requires $2^n-1$ wavelets, where $n$ is the dimension of the space. For example a two-dimensional MRA can be implemented using 2-D scaling and wavelet functions built up using the tensor-product approach \citep{Meyer1992}. In this case three wavelet functions are required to  extract fine features of the field at vertical, horizontal and diagonal orientations. The formulations provided in this paper can be easily extended to two dimensions. However, the computational load of the estimation algorithm will be significant. It is important to investigate methods for choosing basis functions to balance the complexity with the reconstruction capability of the model. \parham{It is also possible to apply other than dyadic subsampling schemes in multiple dimensions, which leads to less wavelet subbands. For example, in 2D, the quincunx scheme is well-known to have better isotropic properties. Unfortunately, in 3D, it is not possible to have a subsampling operator (acting on a Cartesian sampling lattice) that is performing subsampling at equal rate in each dimension at each iteration \citep{VanDeVille2005a}.}
\subsection{Extensions to the Framework} 
Clearly there is, and will always be, a discrepancy between the dynamics of the model and the experimental observation. Nevertheless, the model-based framework proposed in this paper may enable meaningful state tracking and connectivity estimation. The MRAIDE approach is not limited to neural fields; the framework can be applied to modeling other multi-resolution spatiotemporal dynamical systems such as weather systems, ecological systems, and others~\citep{Wikle2002,Xu2005}. The key development is the multi-resolution decomposition forming the state-space model. While the decomposition holds for more sophisticated models, efficiently performing nonlinear smoothing in the large state-space remains a challenge.
